\plaintitle

### 2025-06-02 docs/style/article-body.tex
tabel fix, added MHz for MCU and some small changes

In order to evaluate how efficent our implementation can become, our aim is to try our implementation on other MCUs with varying levels of power and hardware support. Whereas Teensy 4.1 is our primary development platform which we will evaluate closely, we aim to run our implementations on Raspberry Pi Pico 2\footnote{\href{https://datasheets.raspberrypi.com/pico/pico-2-product-brief.pdf}{Raspberry Pi Pico 2 documentation, accessed 2025-03-13}} as well as ESP32-S3\footnote{\href{https://www.espressif.com/sites/default/files/documentation/esp32-s3_datasheet_en.pdf}{ESP32-S3 documentation, accessed 2025-03-13}}. Due to the lower computational power of these MCUs, there may be significant issues in utilizing these weaker models, yet they are significantly cheaper and easier to access. Testing of these will consist solely of running the implementation on these controllers and measuring execution speed and correctness of the output.

@@

Hyncica et. al. \cite{micromeasurements} propose that measuring execution time of algorithms directly via the microcontrollers internal timers (while subtracting the interrupt overhead) provides adequate measurements of the execution speed of an algorithm. An additional advantage is that the same code can be used to measure execution speed on several different microcontrollers, rather than relying on counting CPU cycles (as the process for this may vary greatly between controllers). As we will use fixed-size bitstrings for evaluation, we can then derive the throughput of the algorithm in \(Mbit/s\) as follows:

@@

\begin{table}[ht] \centering \caption{Iteration 1 - Data Structures}
\label{tab:iter1}
\begin{tabularx}{\columnwidth}{|>{\centering\arraybackslash}X|>{\centering\arraybackslash}X|>{\centering\arraybackslash}X|}
\hline \textbf{Bit size} & \textbf{Teensy ($\mu$s)} & \textbf{Pico ($\mu$s)} \\
\hline 64 & 13.1564 & 106.3914 \\ 512 & 788.3139 & 5302.4979 \\ 1024 & 3124.0580
& 21111.2163 \\ \hline \end{tabularx} \end{table}

@@

\begin{table}[ht] \centring \caption{Iteration 2 - Bitshifting}
\label{tab:iter2}
\begin{tabularx}{\columnwidth}{|>{\centering\arraybackslash}X|>{\centering\arraybackslash}X|}
\hline \textbf{Bit size} & \multicolumn{1}{c|}{\textbf{Avg ($\mu s$)}} \\ \hline
64 & 16.4689 \\ 512 & 1006.6255 \\ 1024 & 3996.5972 \\ \hline \end{tabularx}
\end{table}

@@

\begin{table}[H] \centring \caption{Iteration 3 - Batching and Hardware
optimization} \label{tab:iter3}
\begin{tabularx}{\columnwidth}{|>{\centering\arraybackslash}X|>{\centering\arraybackslash}X|}
\hline \textbf{Bit size} & \multicolumn{1}{c|}{\textbf{Avg ($\mu s$)}} \\ \hline
64/64 & 43.0760 \\ 512/512 & 2663.8194 \\ 1024/1024 & 10513.1767 \\ \hline
\end{tabularx} \end{table}

@@

\begin{table}[ht] \centring \caption{Iteration 4 - Loop unrolling}
\label{tab:iter4}
\begin{tabularx}{\columnwidth}{|>{\centering\arraybackslash}X|>{\centering\arraybackslash}X|>{\centering\arraybackslash}X|}
\hline \textbf{Bit size} & \multicolumn{1}{c|}{\textbf{Teensy ($\mu s$)}} &
\multicolumn{1}{c|}{\textbf{Pico ($\mu s$)}} \\ \hline 64 & 9.7017 & 70.7402 \\
512 & 551.6358 & 3978.1928 \\ 1024 & 2195.1979 & 15830.8784 \\ \hline
\end{tabularx} \end{table}

@@

\begin{table}[ht] \centring \caption{Iteration 5 - Removal of vector usage}
\label{tab:iter5}
\begin{tabularx}{\columnwidth}{|>{\centering\arraybackslash}X|>{\centering\arraybackslash}X|>{\centering\arraybackslash}X|}
\hline \textbf{Bit size} & \multicolumn{1}{c|}{\textbf{Teensy ($\mu s$)}} &
\multicolumn{1}{c|}{\textbf{Pico ($\mu s$)}} \\ \hline 64 & 0.0501 & 0.2175 \\
\hline \end{tabularx} \end{table}

@@

\begin{table}[H] \centring \caption{Iteration 6 - Data type exloration}
\label{tab:iter6}
\begin{tabularx}{\columnwidth}{|>{\centering\arraybackslash}X|>{\centering\arraybackslash}X|}
\hline \textbf{Data structure} & \multicolumn{1}{c|}{\textbf{Avg ($\mu s$)}} \\
\hline array & 0.4284 \\ unordered_map & 31.5090 \\ bitset & 0.0474 \\ \hline
\end{tabularx} \end{table}

>>>>

In order to evaluate how efficent our implementation can become, our aim is to try our implementation on other MCU with different level of power and hardware support. Whereas Teensy 4.1, operating at \texttt{600\ MHz}, is our primary development platform which we will evaluate closely, we aim to run our implementations on the Raspberry Pi Pico 2\footnote{\href{https://datasheets.raspberrypi.com/pico/pico-2-product-brief.pdf}{Raspberry Pi Pico 2 documentation, accessed 2025-03-13}}, operating at \texttt{150\ MHz}. Due to the lower computational power of this MCU, there may be significant issues in utilizing this weaker model, yet it is significantly cheaper and easier to access. Testing will consist solely of running the implementation on these controllers and measuring execution speed and correctness of the output.

@@

Hyncica et. al. \cite{micromeasurements} propose that measuring execution time of algorithms directly via the microcontrollers internal timers (while subtracting the interrupt overhead) provides adequate measurements of the execution speed of an algorithm. An additional advantage is that the same code can be used to measure execution speed on several different microcontrollers, rather than relying on counting CPU cycles (as the process for this may vary greatly between controllers).

In our implementation, execution time for each extraction was measured using Arduino's built-in \texttt{micros()}\footnote{\href{https://docs.arduino.cc/language-reference/en/functions/time/micros/}{Documentation for \texttt{micros()}, accessed 2025-05-02.}} function. This function offers a convenient and platform-independent method for recording elapsed time with microsecond precision, enabling consistent performance evaluation across different iterations.

As we will use fixed-size bitstrings for evaluation, we can then derive the throughput of the algorithm in \(Mbit/s\) as follows:

@@

\begin{table}[ht] \centring
\begin{tabularx}{\columnwidth}{|>{\centering\arraybackslash}X|>{\centering\arraybackslash}X|>{\centering\arraybackslash}X|}
\hline \textbf{Bit size} & \textbf{Teensy ($\mu$s)} & \textbf{Pico ($\mu$s)} \\
\hline 64 & 13.1564 & 106.3914 \\ 512 & 788.3139 & 5302.4979 \\ 1024 & 3124.0580
& 21111.2163 \\ \hline \end{tabularx} \caption{Iteration 1 - Data Structures}
\label{tab:iter1} \end{table}

@@

\begin{table}[ht] \centring
\begin{tabularx}{\columnwidth}{|>{\centering\arraybackslash}X|>{\centering\arraybackslash}X|}
\hline \textbf{Bit size} & \multicolumn{1}{c|}{\textbf{Teensy ($\mu s$)}} \\
\hline 64 & 16.4689 \\ 512 & 1006.6255 \\ 1024 & 3996.5972 \\ \hline
\end{tabularx} \caption{Iteration 2 - Bitshifting} \label{tab:iter2} \end{table}

@@

\begin{table}[H] \centring
\begin{tabularx}{\columnwidth}{|>{\centering\arraybackslash}X|>{\centering\arraybackslash}X|}
\hline \textbf{Bit size} & \multicolumn{1}{c|}{\textbf{Teensy ($\mu s$)}} \\
\hline 64/64 & 43.0760 \\ 512/512 & 2663.8194 \\ 1024/1024 & 10513.1767 \\
\hline \end{tabularx} \caption{Iteration 3 - Batching and Hardware optimization}
\label{tab:iter3} \end{table}

@@

\begin{table}[ht] \centring
\begin{tabularx}{\columnwidth}{|>{\centering\arraybackslash}X|>{\centering\arraybackslash}X|>{\centering\arraybackslash}X|}
\hline \textbf{Bit size} & \multicolumn{1}{c|}{\textbf{Teensy ($\mu s$)}} &
\multicolumn{1}{c|}{\textbf{Pico ($\mu s$)}} \\ \hline 64 & 9.7017 & 70.7402 \\
512 & 551.6358 & 3978.1928 \\ 1024 & 2195.1979 & 15830.8784 \\ \hline
\end{tabularx} \caption{Iteration 4 - Loop unrolling} \label{tab:iter4}
\end{table}

@@

\begin{table}[ht] \centring
\begin{tabularx}{\columnwidth}{|>{\centering\arraybackslash}X|>{\centering\arraybackslash}X|>{\centering\arraybackslash}X|}
\hline \textbf{Bit size} & \multicolumn{1}{c|}{\textbf{Teensy ($\mu s$)}} &
\multicolumn{1}{c|}{\textbf{Pico ($\mu s$)}} \\ \hline 64 & 0.0501 & 0.2175 \\
\hline \end{tabularx} \caption{Iteration 5 - Removal of vector usage}
\label{tab:iter5} \end{table}

@@

\begin{table}[H] \centring
\begin{tabularx}{\columnwidth}{|>{\centering\arraybackslash}X|>{\centering\arraybackslash}X|}
\hline \textbf{Data structure} & \multicolumn{1}{c|}{\textbf{Teensy ($\mu s$)}}
\\ \hline array & 0.4284 \\ unordered_map & 31.5090 \\ bitset & 0.0474 \\ \hline
\end{tabularx} \caption{Iteration 6 - Data type exloration} \label{tab:iter6}
\end{table}
